---
title: "tidytext In_Class Assignment"
author: "Include Student Name Here"
date: '2024-02-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The questions in this markdown must be answered using the tidyr and dplyr packages' functions. 
#### The tidyverse functionality is greatly enhanced using the pipes (%>% from magrittr)
#### Load the required packages before starting the assignment. 
#### Questions 1 through 11 are related to tidyr (with one exception of using a base function - "replace")
#### Questions 12 through 11 are related to dplyr (with one exception of using a base function - "replace")

```{r}
library(tidyr)
library(dplyr)
library(lubridate)
library(tibble)
library(ggplot2)
```


```{r Q1. religious_income}
# Take a look at the religious income dataset, relig_income and answer if the dataset is a tidy dataset. 
#NO

# If not, why is it not a tidy dataset?

# THere are ranges when there should be one column for income

# Tidy the dataset using pivot_longer transformation.

tidyData <- relig_income %>%
  pivot_longer(cols = -religion,
               names_to = "income_range",
               values_to = "income")

```


``` {r Q2. transform_subset_of_columns}
student_scores <- tribble(
  ~name, ~bus101, ~bus201, ~eco101, ~fin101, 
  "Billy", 90, 96, 70, 76, 
  "Suzy", 82, 76, 78, 86,
  "Lionel", 68, 56, 79, 82, 
  "Jenny", 72, 96, 71, 90,
)

# Merge the Business courses into a single column with the name "Business Courses" and the corresponding values from their respective column names. While saving the values, only use numeric course numbers 101 and 201.

# Use names_prefix to supress character values from the column names

student_scores %>%
  pivot_longer(cols = c(bus101, bus201),
               names_to = "Business Courses",
               values_to = "Scores",
               names_prefix = "bus")

```


``` {r Q4. Split the Business courses and course numbers}
# Use the names pattern function to split required values. The value "bus101" should be split such that "bus" appears in the "major" column and "101" appears in the "course id" column.
# Use names_pattern to specify how each column should be split into the specified number of columns.
# Use names_to to provide the names of the new columns. 

# names_Pattern(): {} = number of characters []= range of what we are looking for
student_scores %>%
  pivot_longer(cols = -name,
               names_pattern = "([a-z]{3})([0-9]*)",
               names_to = c("major", "course_id"),
               values_to = "Scores")

  



```



``` {r Q5. }
# Take a look at each of these tables and answer which table is tidy and which tables are not tidy. 
# table1 tidy or not?
# table2 tidy or not?
# table3 tidy or not?
# table4a tidy or not?
# table4b tidy or not?

#'table1, table2, table3, table4a, table4b, and table5 all display the number of  TB cases documented by the World Health Organization in Afghanistan, Brazil, and China between 1999 and 2000. The data contains values associated with four variables (country, year, cases, and population), but each table organizes the values in a different layout. 
#'The data is a subset of the data contained in the World Health Organization Global Tuberculosis Report
#'
# table1 is tidy.
# table2 is tidy
# table3 is not tidy based on the rate 
# table4a is not tidy becasue the year feature is captured in two different columns/variables.
# table4b is not tidy becasue the year feature is captured in two different columns/variables.

```



``` {r Q6. Split the data from type into two columns, with cases and population as column names}
# Using table2, split the "type" column into two columns, with cases and population as column names. User pivot_wider() function
table2 %>%
  pivot_wider(names_from = type,
              values_from = count)

```


``` {r Q7. Separate the rate column into two columns, cases and population}
# Using table3, separate the rate column into two columns, cases and population. Use separate() function 

table3 %>%
  separate(rate, into = c("cases", "population"), 
           sep = "/")

```


``` {r Q8. combine the day, month, and year into DOB column}
student_details <- tribble(
  ~name, ~day, ~month, ~year, ~"pass/fail", 
  "Billy", 15, 04, 96, "Pass", 
  "Suzy", 21, 10, 98, "Fail",
  "Lionel", 27, 09, 97, "Pass", 
  "Jenny", 01, 04, 99, "Pass"
)
student_details %>%
  unite(col="DOB",
        month, day, year, sep="-")

```


``` {r Q9. Populate all columns}
# Use the complete() function of tidyr to insert rows with all possible combinations of name and course columns. 
student_scores <- tribble( 
  ~name, ~course, ~score, 
  "Billy", "bus101", 90,
  "Billy", "fin101", 76,
  "Suzy", "bus101", 82,
  "Suzy", "bus201", 76,
  "Suzy", "eco101", 78,
  "Suzy", "fin101", 86,
  "Lionel", "bus101", 68,
  "Lionel", "bus201", 56,
  "Lionel", "eco101", 79,
  "Lionel", "fin101", NA )

student_scores %>%
  complete(name, course)

```


``` {r Q10. Replace the negative numbers with 0}
# Find the indexes or row numbers of sleep dataset that has negative numbers in extra column 
# Make sure that the indexes are of integer type (Numeric)

# Use base package replace function to the negative numbers with 0
sleep$extra <- replace(sleep$extra, sleep$extra < 0, 0)
sleep



```


``` {r Q11. use replace_na() to replace NA with "Not Registered"}
student_scores <- tribble(
  ~name, ~course, ~score,

  "Billy",  "bus101", 90,
  "Billy",  "bus201", NA,
  "Billy", "eco101", NA,
  "Billy", "fin101", 76,
  "Lionel", "bus101", 68,
  "Lionel", "bus201", 56,
  "Lionel", "eco101", 79,
  "Lionel", "fin101", NA,
  "Suzy", "bus101", 82,
  "Suzy", "bus201", 76,
  "Suzy", "eco101", 78,
  "Suzy", "fin101", 86
)
student_scores %>%
  replace_na(list(score=00))

``` 

#### dplyr functions 

```{r Q12. inner_join and semi_join - What is the difference?}

superheroes <- tibble::tribble(
       ~name, ~alignment,  ~gender,          ~publisher,
   "Magneto",      "bad",   "male",            "Marvel",
     "Storm",     "good", "female",            "Marvel",
  "Mystique",      "bad", "female",            "Marvel",
    "Batman",     "good",   "male",                "DC",
     "Joker",      "bad",   "male",                "DC",
  "Catwoman",      "bad", "female",                "DC",
   "Hellboy",     "good",   "male", "Dark Horse Comics"
  )

publishers <- tibble::tribble(
  ~publisher, ~yr_founded, ~friend,
        "DC",       1934L, "jon",
    "Marvel",       1939L, "Hell",
     "Image",       1992L, "Merc"
  )

# Use inner_join to join superheroes and publishers on publisher column.
superheroes %>% 
  inner_join(publishers) %>%
  select(name, alignment, friend)

# Use semi_join to join superheroes and publishers on publisher column.
superheroes %>% 
  semi_join(publishers)


# What is the difference between these two types of joins? 

#Semi ignores the second column (yr_founded). only joins the first column.

```


```{r Q13. Filter rows by one or more column conditions}
# Display the diamond count by cut and color. Include only the rows with a diamond count greater than 300. Use the filter function instead of [] operation. 

# CAUTION! Not price as below:
#diamonds[diamonds$price > 300, ]

diamonds %>% 
  count(cut, color) %>%
  filter(n > 300)




```

```{r just_for_fun_with_filter}
#install.packages("nycflights13")
library(nycflights13) # Requires package installation
 
airports %>% filter(faa %in% c("PDX", "DCA", "SEA", "JAC", "SAT", "ORD", "MIA") ) %>% 
  ggplot() +
     borders("state", colour = "blue") +
     geom_point(aes(lon, lat)) + 
     geom_text(aes(lon, lat, label = name), color = "brown", vjust = 1.25)+
     coord_quickmap()

```

``` {r Q14. Slice the dataset diamonds}
# Using slice function fetch the rows from 10th row to 15th row

# Uncomment the code below to complete the response 
diamonds %>%
  slice(10:15)

```


``` {r Q15. Using mutate, normalize diamonds price}
# Using mutate function create a normalized price value. Create a new column price_normal
# price_normal is (price - mean(price))/ standard_deviation(price)
# No need to save the output dataset to a new diamonds dataset 
# Verify that the new column was created correctly by including the following line of code at the end of the pipe operation
# filter(price == 3932 | price == 3935)

diamonds %>% 
  mutate(price_normal = (price - mean(price))/sd(price)) %>%
  filter(price == 3932 | price == 3935)

```



#### The scoped variants of mutate() and transmute() make it easy to apply the same transformation to multiple variables. There are three variants:

- _all affects every variable    

- _at affects variables selected with a character vector or vars()    

- _if affects variables selected with a predicate function:    
``` {r Q16. using mutate_at add $10 to the price}
# using mutate_at add $10 to the price
# This is the syntax of mutate_at
# mutate_at(.tbl, .vars, .funs, ..., .cols = NULL) 

diamonds %>% 
  mutate_at(c("price", "depth"), function(x) x + 10)

```


``` {r Q17. Use mutate across to add $10 to the price}
# using mutate across, add $10 to the price. mutate across is the preferred method. 
# use mutate(across( .. rest all should be the same
# mutate(across(.cols = everything(), .fns = NULL, ..., .names = NULL) 

diamonds %>% 
  mutate(across(
    "price", .fns=function(x) x + 10
  ))

```


``` {r Q18. using mutate_if increase price by $10}
# using mutate_if increase price by $10
# mutate_if(.tbl, .predicate, .funs, ...)

diamonds %>%
  mutate_if(colnames(.) %in% c("price", "depth"), function(x) x + 10)

```


``` {r Q19. using mutate create a discounted_price}
# using mutate function add a new column to the diamonds dataset called discounted_price to capture discounted price depending on the cut of the diamond.  
# if "Fair" - price*.90, if "Good" - price*.85, if "Very Good" - price*.80, if "Premium" - price*.75, if "Ideal" - price*.75
diamonds %>% 
  mutate(discounted_price = case_when(cut == "Fair" ~ 0.9*price,
                                      cut == "Good" ~ 0.85*price,
                                      cut == "Very Good" ~ 0.8*price,
                                      cut == "Premium" ~ 0.75*price,
                                      TRUE ~ price))


```



``` {r Q20. Sort the diamonds dataset by cut and select 3 columns}
# sort the diamonds dataset by cut column and select, price, cut and depth columns 
diamonds %>% 
  arrange(desc(cut)) %>%
  select(price, cut, depth)
```


``` {r Q21. summarize the diamonds dataset and calculate mean_price, mean_depth, variance_x}
# Using the diamonds dataset and summarize function, calculate the mean_price, mean_depth, and variance_x (using x variable) 

diamonds %>% 
  group_by(cut) %>%
  summarise(mean_price = mean(price),
            mean_depth = mean(depth))

```


``` {r Q22. Create a cumulative hourly increase of 0.001 in the average crypto price. }
# Complete the code below to generate a dataset containing the datetime and price of a fictitious cryptocoin with a consistent hourly increase of 0.001 in value.)
library(lubridate)
march_hours <- ymd_hms("2023-03-01 00:00:00") + seq(0:200)*60*60 # Some 200 random hours

# Uncomment the line of code below to complete a mutate & a cumsum function and create a variable that stores the cumulative price increase. Utilize a rnorm with a mean of 0.001 and a standard deviation of 0.01

crypto_price <- as_tibble(march_hours) %>% 
  mutate(price = 0.07 + 
           cumsum(rnorm(n(), mean = 0.001, sd = 0.01)))

# Create a plot to plot the daily crypto price chart of this pseudo coin. 
plot(crypto_price, 
     type = 'l')

```

``` {r 23. create a union of the diamonds1 and diamonds2 datasets}
# create a union of the diamonds1 and diamonds2 datasets
# Look at the cheatsheet for understanding these functions. Also, please pay attention to the intersect function you are using. 
diamonds1 <- diamonds[diamonds$depth < 50,]
nrow(diamonds1)
diamonds2 <- diamonds[diamonds$depth < 55,]
nrow(diamonds2)
nrow(union_all(diamonds1, diamonds2))


```

``` {r 24. create a intersection of the diamonds1 and diamonds2 datasets}
# create a intersection of the diamonds1 and diamonds2 datasets
nrow(dplyr::intersect(diamonds1, diamonds2))
```


``` {r 25. use setdiff function of dplyr to find the rows that differ diamonds1 and diamonds2 datasets}
# use setdiff function of dplyr to find the rows that differ diamonds1 and diamonds2 datasets
nrow(dplyr::setdiff(diamonds2, diamonds1))
```
